---
title: "Exercse 3"
output:
  html_document:
    df_print: paged
---

# Task 1: Segmentation
Load fun packages.
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggpubr)
library(SimilarityMeasures)
theme_set(
  theme_classic()
)
```

Import caro data
```{r}
caro <- read_csv("0_rawdata/caro60.csv")
caro_original <- caro
```
## Specify a temporal window
Our sampling interval is 1 minute - if we take a temporal window of 6 minutes we use a window size of 6 positions. Therefore that's 3 plus and 3 minus.
```{r}
caro %>% 
  mutate(
    nMinus3 = sqrt((lag(E,3)-E)^2 + (lag(N,3)-N)^2),
    nMinus2 = sqrt((lag(E,2)-E)^2+(lag(N,2)-N)^2),  
    nMinus1 = sqrt((lag(E,1)-E)^2+(lag(N,1)-N)^2),
    nPlus1  = sqrt((E-lead(E,1))^2+(N-lead(N,1))^2), 
    nPlus2  = sqrt((E-lead(E,2))^2+(N-lead(N,2))^2),  
    nPlus3  = sqrt((E-lead(E,3))^2+(N-lead(N,3))^2)  
) -> caro

caro %>% 
  rowwise() %>% 
  mutate(
    stepMean = mean(c(nMinus3, nMinus2, nMinus1, nPlus1, nPlus2, nPlus3))
  ) %>% 
  ungroup() -> caro
```


# Task 2: Specify and apply threshold *d*
Lets explore with a boxplot, histogram and summary.
```{r}
ggplot(caro) +
  geom_boxplot(aes(y = stepMean), outlier.shape = NA)  +
  scale_y_continuous(limits = c(0, 20))

ggplot(caro) +
  geom_histogram(aes(stepMean), bins = 60) + geom_vline(xintercept = 4, color = "red")

summary(caro$stepMean)

ggplot(caro) +
  geom_line(aes(x=DatetimeUTC, y=stepMean)) +
  geom_hline(yintercept = 4, color = "red")
```
If we assume that the animal rests more than moving a value of d =< is fitting. The line graph highlights that movement below that point largely seems to be noise.

## Remove static points
Considering that a threshold value of 4 is not far off the mean (3.9) we can just use the mean.
```{r}
caro %>% 
  ungroup() %>% 
  mutate(static = stepMean < mean(stepMean, na.rm = TRUE)) -> caro

caro %>% 
  filter(!static) -> caro_filtered
  ggplot(data = caro_filtered, aes(x = E, y = N)) +
    geom_path() +
    geom_point() +
    coord_fixed()
```
# Task 3 Visualise segmented trajectories
```{r}
ggplot(caro, aes(E, N)) +
  geom_path() +
  geom_point(aes(color = static)) +
  coord_equal()
```
# Task 4: Segment based analysis
Functions to create unique IDs
```{r}
rle_id <- function(vec){
  x <- rle(vec)$lengths
  as.factor(rep(seq_along(x), times=x))
  }
```

Assign unique IDs to Caro and visualise all the segments.
```{r}
caro <- caro %>%
  mutate(segment_id = rle_id(static))

ggplot(caro, aes(E, N, color = segment_id)) +
  geom_path() +
  geom_point() +
  coord_equal()

caro
```
Now lets calculate the duration of each segment and remove the shorter segments (< 5 minutes)
```{r}
caro %>% 
  mutate(timelag = as.integer(difftime(lead(DatetimeUTC),DatetimeUTC, units = "mins"))) %>%
  group_by(segment_id) %>% 
  summarise(total_time = sum(timelag)) -> segment_duration

left_join(caro, segment_duration, by = "segment_id") -> caro
```


Makes plots
```{r}
caro %>% 
  filter(!static) %>% 
  filter(total_time >= 5) %>% 
  ggplot(., aes(E, N, color = segment_id)) +
  geom_path() +
  geom_point() +
  coord_equal() + labs(title="Removed segments shorter than 5 minutes")
```
# Task 5: Similarity measures
Import data
```{r}
pd <- read_csv("0_rawdata/pedestrian.csv")
```

Explore the trajectories
```{r}
pd
ggplot(pd, aes(E, N, color = DatetimeUTC)) +
  geom_path() + geom_point() + facet_wrap(~TrajID)

summary(pd)

pd %>% 
  group_by(TrajID) %>% 
  summarise(n = n())
```
Trajectory 3 moved in reverse. Trajectory 4 has a differing shape and 5 has an outlier.
Point density between the trajectories varies and highlights differences in speed.

# Task 6: Calculate similarity

First we need to get matrices!

	
An m x n matrix containing trajectory1. 
Here m is the number of points and n is the dimension of the points.

What does dimension mean?
```{r}
trajectories <- seq(2,5,1)
df <- NULL

for (i in trajectories) {
  pd %>% filter(TrajID == 1) %>% select(E, N) %>%  as.matrix() -> trj_1
  pd %>% filter(TrajID == i) %>% select(E, N) %>%  as.matrix() -> trj_2
  DTW(trj_1, trj_2) -> dtw_value
  EditDist(trj_1, trj_2) -> edit_dist_value
  Frechet(trj_1, trj_2) -> frechet_value
  LCSS(trj_1, trj_2, errorMarg = 15) -> lcss_value
  df <- rbind(df, data.frame(i, dtw_value, edit_dist_value, frechet_value, lcss_value))

}


```

Lets visualise the 3 methods
```{r}
df %>% 
  rename(DTW = dtw_value, EditDist = edit_dist_value, Frechet = frechet_value, LCSS = lcss_value) %>% 
  pivot_longer(2:5, names_to ="methods", values_to = "computed_similarities")%>% 
  ggplot(., aes(x=as.factor(i), y=computed_similarities, fill = as.factor(i))) +
    geom_col() + facet_wrap(~methods, scales = "free_y") + theme(legend.position="none") +
  xlab("Trajectories") + ylab("Computed similarity scores")
```

